Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2024-08-08 paper/main.tex
Updates

\begin{abstract}
Causal discovery has received significant attention in the Directed Acyclic
Graphs (DAGs) literature, leading to the development of numerous automated
algorithms for learning DAGs from data. Despite these advancements, the
application of these algorithms in practical domains remains limited, with
researchers often preferring to manually construct DAGs based on their expert
knowledge. Even when using these algorithms, researchers must rely on their
expertise to correct mistakes or orient edges before the models can be used for
further analysis. Given this reliance on manual model construction and
modification, we propose a method inspired by specification search in
Structural Equation Models (SEMs) to assist researchers in iteratively
modifying DAGs. Our method employs implied Conditional Independence (CI) based
model testing, combined with a novel measure of partial association for mixed
data, to rank violations of implied CIs of the model in the data. This ranking
helps users to prioritize pair of variables between which to add or remove
edges. Users can then decide to remove or add edges while determining the
correct orientation of the edge based on their expertise. Empirically, we
demonstrate that a greedy approach (i.e., fixing the worst violation first)
achieves performance comparable to automated algorithms, provided the user can
correctly determine the edge direction in $ 1 $ out of $ 3 $ edge additions.
Additionally, we show that in the absence of an expert, a Large Language Model
(LLM) can potentially be used to determine edge orientations. 
\end{abstract}

>>>>

\begin{abstract}
Causal discovery has received significant attention in the Directed Acyclic
Graphs (DAGs) literature, leading to the development of numerous automated
algorithms for learning DAGs from data. Despite these advancements, the
application of these algorithms in practical domains remains limited, with
researchers often preferring to manually construct DAGs based on their expert
knowledge. Even when using these algorithms, researchers must rely on their
expertise to correct mistakes or orient edges before the models can be used for
further analysis. Given this reliance on manual model construction and
modification, we propose a method inspired by specification search in
Structural Equation Models (SEMs) to assist researchers in iteratively
modifying DAGs. Our method employs implied Conditional Independence (CI) based
model testing, combined with a novel measure of partial association for mixed
data, to rank violations of implied CIs of the model in the data. This ranking
helps users to prioritize pair of variables between which to add or remove
edges. Users can then decide to remove or add edges while determining the
correct orientation of the edge based on their expertise. Empirically, we
demonstrate that a greedy approach (i.e., fixing the worst violation first)
achieves performance comparable to automated algorithms, provided the user can
correctly determine the edge orientation in $ 1 $ out of $ 3 $ edge additions.
Additionally, we show that in the absence of an expert, a Large Language Model
(LLM) can potentially be used to determine edge orientations. 
\end{abstract}
